---
subtitle: "TMA4268 Statistical Learning V2018"
title: "Compulsory exercise 1: Group 21"
author: "Jørgen Opheim, Ole-Andreas Sandnes and Sander Coates"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: #3rd letter intentation hierarchy
  # prettydoc::html_pretty:
  #   theme: tactile
  #   highlight: github
  pdf_document:
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Take a look at the cheat sheets for R Markdown here: File > Help > Cheatsheets > R Markdown Cheat Sheet in RStudio, or here
<http://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf>

or the lessons: <http://rmarkdown.rstudio.com/lesson-1.html>

# Problem 1 - Core concepts in statistical learning [2 points]

## a) Training and test MSE [1 point]

* *Figure 2* shows that variance is reduced for increased values of $K$, but at the cost of increased bias.
* A low value of $K$ gives the most flexible fit.

* As expected from lower flexibility for higher $K$, the training MSE increases with $K$. This is due to the reduced fitting of the model to the specific data. However, when introducing the test data, overfitting by a too flexible model leads to increased test MSE for the lowest values of $K$. This suggest a slightly less flexible and thus less biased model (more on the bias-variance trade-off later) is better.
* By observation, it seems that $K = 3$ gives the lowest test MSE, and hence is the best choice of K for modelling $f(x)$ based on observed values $y = f(x) + \epsilon$.

## b) Bias-variance trade-off [1 point]

* \textcolor{blue}{Explain how that is done. Hint: this is what the $M$ repeated training data sets are used for.}
  The variance is calculated by use of R's own function ```var``` over all experiments $M$ for a given $x$ and $K$. The squared bias is then found by squaring the difference between the mean over all $M$ experiments for a given $x$ and $K$ and the true value of $y$ for that particular $x$ (equal for all values of $K$). \textcolor{red}{Add some formulae?}
  
* Focus on Figure 4. As the flexibility of the model increases ($K$ decreases), what happens with
    + the squared bias,  
    + the variance, and  
    + the irreducible error?
* What would you recommend is the optimal value of $K$? Is this in agreement with what you found in a)?

# Problem 2 - Linear regression [4 points]

Here you see an R chunk that is evaluated (when knitting) and code is displayed.

```{r,echo=TRUE,eval=FALSE}
library(ggplot2)
data = read.table("https://www.math.ntnu.no/emner/TMA4268/2018v/data/SYSBPreg3uid.txt")
dim(data)
colnames(data)
modelA=lm(-1/sqrt(SYSBP) ~ .,data = data)
summary(modelA)
```

## a) Understanding model output [1 point]
## b) Model fit [1 point]
## c) Confidence interval and hypothesis test [1 points]
## d) Prediction [1 point]