---
subtitle: "TMA4268 Statistical Learning V2018"
title: "Compulsory exercise 1: Group 21"
author: "Jørgen Opheim, Ole-Andreas Sandnes and Sander Coates"
date: "[Date]"
output: #3rd letter intentation hierarchy
  # prettydoc::html_pretty:
  #   theme: tactile
  #   highlight: github
  pdf_document:
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Take a look at the cheat sheets for R Markdown here: File > Help > Cheatsheets > R Markdown Cheat Sheet in RStudio, or here
<http://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf>

or the lessons: <http://rmarkdown.rstudio.com/lesson-1.html>

# Problem 1 - Core concepts in statistical learning [2 points]

## a) Training and test MSE [1 point]
Here you write your smart answer. You may write latex $Y=f(x)+\varepsilon$.

Items with

* item1
* item2

## b) Bias-variance trade-off [1 point]

* Explain how that is done. Hint: this is what the $M$ repeated training data sets are used for. 
* Focus on Figure 4. As the flexibility of the model increases ($K$ decreases), what happens with
    + the squared bias,  
    + the variance, and  
    + the irreducible error?
* What would you recommend is the optimal value of $K$? Is this in agreement with what you found in a)?

# Problem 2 - Linear regression [4 points]

Here you see an R chunk that is evaluated (when knitting) and code is displayed.

```{r,echo=TRUE,eval=FALSE}
library(ggplot2)
data = read.table("https://www.math.ntnu.no/emner/TMA4268/2018v/data/SYSBPreg3uid.txt")
dim(data)
colnames(data)
modelA=lm(-1/sqrt(SYSBP) ~ .,data = data)
summary(modelA)
```

## a) Understanding model output [1 point]
## b) Model fit [1 point]
## c) Confidence interval and hypothesis test [1 points]
## d) Prediction [1 point]